---
title: "Covid91 vaccine study Final2023"
author: "Cyndhell Anne Pilarca"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$: The Covid19 vaccine (drug) is not effective among subgroups users.   
$H_A$: The Covid19 vaccine (drug) is effective among subgroups users.


# Males

$H_0$:The vaccine is not effective in preventing Covid-19 in males   
$H_A$: The vaccine is effective in preventing Covid-19 in males

Data for this study came from a FinalData file, provided by the professor.

## Methods

The response method used is Cat~Cat. Data used was based on data from FinalData. We have to subset the data to include male.   
Graphical display: bar chart, numerical display and inferential results

## Results

```{r}
maleData = filter (FinalData,sex == "M")
```

## Descriptive Results

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=maleData)
barchartGC(~infected + treatment,data=maleData, type="percent")
```

According to the graph, there was a difference between 30% with Covid-19 and 70% with placebo among the guys who received the vaccination and those who received Covid-19. There was little difference among those who received the vaccinations and were "OK."

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment, data= maleData)
rowPerc(table1)
colPerc(table1)
```

The table demonstrates that there was a 2 times difference among those who received the vaccinations and experienced Covid-19. 70% of those who took the placebo did not get the disease, compared to 30% of those who got the vaccination. There was not a noticeable distinction among those who received vaccinations and were OK.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for male is 0.0000000000001098 the p-value is too small that's why we failed to reject the null hypothesis. The vaccine is not effective in preventing Covid-19 among males. Based on the Fisher's Exact test the P-value is 0.00000000000004631 meaning there is a difference in getting Covid-19 and being a male. If you are a males, your chances of acquiring Covid-19 are the same whether you take the placebo or Covid-19.


# Females

$H_0$:The vaccine is not effective in preventing Covid-19 in Females.      
$H_A$:  The vaccine is effective in preventing Covid-19 in Females.    
Data for this study came from a FinalData file, provided by the professor.

## Methods

The response method used is Cat~Cat. Data used was based on data from FinalData. We have to subset the data to include female.        
Graphical display: bar chart, numerical display and inferential results numerical display

## Results

```{r}
femaleData = filter (FinalData,sex == "F")
```

## Descriptive Results

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=femaleData)
barchartGC(~infected + treatment,data=femaleData, type="percent")
```

According to the graph, among those who received Covid-19, there was a difference between 40% Covid-19 and 60% placebo among the females who received the vaccination. The Female who received the vaccination and were OK showed no noticeable difference.

### Numerical Descriptive Results

```{r}
table2 <- xtabs(~infected + treatment, data=femaleData)
rowPerc(table2)
colPerc(table2)
```

According to the table, there was a 20% difference in vaccination rates among females who received the Covid-19 vaccine. The female who recieve the vaccine (40%) and the placebo (60%). There is no difference among those who received the vaccination.


## Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The Chi-squared p-value is 0.002732 the p-value is too small, so the vaccine is not effective in preventing covid-19 among females. Based on the Fisher's Exact test the p-value is 0.0002162. We failed to reject the null hypothesis, because the p-value is too small.


# LGBTQ

$H_0$: The vaccine is not effective in preventing Covid in LGBTQ   
$H_A$: The vaccine is effective in preventing Covid in LGBTQ    
Data for this study came from a FinalData file, provided by the professor.

## Methods

The response method used is Cat~Cat. Data used was based on data from FinalData. We have to subset the data to include LGBTQ.    
Graphical display: bar chart, numerical display and inferential results numerical display

## Results

```{r}
LGBTQ = filter (FinalData,LGBTQ == "gay")
```

## Descriptive Results
### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")
```

According to the graph, among individuals who received the vaccination and had Covid-19, there was an 80% Covid-19 compared. 20% placebo difference. There was no apparent variance among those who received vaccinations and were OK.


### Numerical Descriptive Results

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table3)
colPerc(table3)
```

According to the table, there was a 40% difference in individuals who contracted Covid-19 among those who received the vaccination. About 80% of those who received the placebo had a 20% chance of not becoming sick. There was no discernible difference among those who received vaccinations and were OK.

## Inferential Results


```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

The Chi-squared value is 0.00000006668 is too small, so vaccine is not effective in preventing spread of vaccine among LGBTQ. Based on the Fisher's Exact test P-value are 0.00000001029. We fail to reject the null hypothesis. If you take the placebo or covid you are in the same likely hood of receiving covid if you are a LGBTQ.


# Druggies

$H_0$: The vaccine is not effective in preventing Covid in Druggies    
$H_A$: The vaccine is effective in preventing Covid in Druggies    
Data for this study came from a FinalData file, provided by the professor.

## Methods

The response method used is Cat~Cat used was based on data from FinalData. We have to subset the data to include Druggies.   
Graphical display: bar chart, numerical display and inferential results numerical display

## Results

```{r}
DrugUser = filter (FinalData, DrugUser == "yes")
```

## Descriptive Results

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")
```

The graph shows among those who had Covid-19, the Druggies who did get vaccinated there was a difference between 88% with Covid-19 and about 10% with placebo. 

### Numerical Descriptive Results

```{r}
table4 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table4)
colPerc(table4)
```

The table shows that among the Druggies who took the vaccine that got Covid-19 there was a difference. The Druggies that took the placebo with 12% chance of not being infected and 88% for the Druggies. Among the people that were OK and took the vaccine there was no significant diligence.


## Inferential Results

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The Chi-squared value is p-value 0.000000002842 is too small, so vaccine is not effective in preventing spread of Covid-19 among other people. Based on the Fisher's Exact test the p-value is 0.000000000232. We fail to reject the null hypothesis. The information provided shows that the vaccine does not prevent Covid-19.


# Overall Results and Conclusions

In the original question we asked if the vaccine was able to prevent Covid-19 among these subgroups: males, females, LGBTQ and drug users. We have found that the vaccine is ineffective at preventing Covid-19 in those subgroups, therefore we fail to reject the null hypothesis.

We aimed for to know if Covid-19 was successfully prevented by the vaccinations in the male subgroup. Our findings ultimately lead us to draw the conclusion that the vaccination is ineffective in protecting males against COVID-19. We fail to reject the null hypothesis for this subgroup. We fail to reject the null hypothesis since the vaccination was unable to prevent Covid-19 in the subgroup of females when the same question was presented. The results of our research show that the vaccination did not successfully prevent Covid-19, and since the LGBTQ population was also included in the study, we are fail to reject the null hypothesis. Like as well in the drug users asked the same questions, and because our findings show that the vaccinations did not prevent Covid-19, we failed to reject the null hypothesis.

